\begin{abstract}
    In order to combine the extensive developments made in SQL queries over the years with the benefits of graph queries, the striking SQL/Property Graph Queries (abbr.~SQL/PGQ) is proposed.
    However, since the SQL/PGQ queries contain both relational and graph queries, it is challenging for existing optimizers in relational databases to optimize these queries.
    In this paper, we propose a matching operator for graph pattern matching to correspond to graph queries in SQL/PGQ queries, and abstract the SQL/PGQ queries into SPJM (Select-Project-Join-Match) queries. 
    Then, we propose a converged graph relational optimization framework named relgo to optimize SPJM queries.
    Relgo optimizes matching operators by decomposing patterns, which utilizes graph-specific optimization techniques.
    It is proved that this graph-aware method is exponentially faster than graph-agnostic method in optimization.
    Additionally, relgo regards the physical implementation of matching operators as a specialized implementation of scan operators, termed ScanMatchTable.
    Since the output of ScanMatchTable is a relation, the results can be optimized together with other relational operators.
    Moreover, a new physical implementation of join operators named \expandintersect has been proposed to further improve the efficiency of plan execution.
    Various experiments have been conducted to show the efficiency of relgo.
    End-to-end experiments demonstrate the superior performance of relgo.
    In detail, the execution time of plans generated by relgo can be more than 20$\times$ shorter than that of the plan generated by GrainDB.
    Besides, the experimental results of ablation studies illustrate the efficiency of \expandintersect.

\end{abstract}
